The Strucplot Framework---Visualizing Multi-way Contingency Tables

### 2005-10-16 trunk/vcd/inst/doc/shading.Rnw
revert class system to minimal level
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@471 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<<shading2>>=
shading2_fun <- function(res) gpar(fill = ifelse(res > 0, "lightblue", "lightsalmon"))
class(shading2_fun) <- c("grapcon","shading")
@

\noindent Note the \code{class} attribute set on \codefun{shading2\_fun}.
To create a mosaic display with binary shading, 
it now suffices to specify the data table along with
\codefun{shading2\_fun}:

@@

<<shading3>>=
shading3a_fun <- function(col = c("lightblue", "lightsalmon")) {
  if (length(col) != 2) stop("Need exactly two colors!")
  ret <- function(res) gpar(fill = ifelse(res > 0, col[1], col[2]))  
  structure(ret, class = c("grapcon","shading"))
}
@

@@

<<shading4>>=
shading3b_fun <- function(observed = NULL, residuals = NULL, expected = NULL, 
    df = NULL, col = c("lightblue", "lightsalmon")) {
  if (length(col) != 2) stop("Need exactly two colors!")
  ret <- function(res) gpar(fill = ifelse(res > 0, col[1], col[2]))  
  structure(ret, class = c("grapcon","shading"))
}
class(shading3b_fun) <- c("grapcon_generator","shading")
@

>>>>

<<shading2>>=
shading2_fun <- function(res) gpar(fill = ifelse(res > 0, "lightblue", "lightsalmon"))
@

\noindent To create a mosaic display with binary shading, 
it now suffices to specify the data table along with
\codefun{shading2\_fun}:

@@

<<shading3>>=
shading3a_fun <- function(col = c("lightblue", "lightsalmon")) {
  if (length(col) != 2) stop("Need exactly two colors!")
  function(res) gpar(fill = ifelse(res > 0, col[1], col[2]))  
}
@

@@

<<shading4>>=
shading3b_fun <- function(observed = NULL, residuals = NULL, expected = NULL, 
    df = NULL, col = c("lightblue", "lightsalmon")) {
  if (length(col) != 2) stop("Need exactly two colors!")
  function(res) gpar(fill = ifelse(res > 0, col[1], col[2]))  
}
class(shading3b_fun) <- "grapcon_generator"
@

### 2005-10-16 trunk/vcd/inst/doc/strucplot.Rnw
revert class system to minimal level
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@471 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The strucplot framework is highly modularized: Figure \ref{fig:struc}
shows the hierarchical relationship between the various components. 
On the lowest level, there are several groups of workhorse and 
parameter functions that directly or indirectly influence the final
appearance of the plot. These `graphical appearance control'
(`grapcon') functions are created by generating functions, allowing
flexible parameterization and extensibility. 
The first part of their names (\code{\emph{group\_foo()}}) 
reflects the group they belong to (strucplot core, labeling,
legend, shading, or spacing). The workhorse functions (created by 
\code{\emph{struc\_foo()}}, 
\code{\emph{labeling\_foo()}}, and \code{\emph{legend\_foo()}}) 
directly produce graphical output (``add ink to the canvas''), whereas
the parameter functions
(created by \code{\emph{spacing\_foo()}} and \code{\emph{shading\_foo()}}) compute
graphical parameters used by the others. The grapcon functions returned by 
\code{\emph{struc\_foo()}} implement the core functionality, creating the tiles and their
content.  On the second level of the framework, a suitable combination
of grapcon functions (or, alternatively, corresponding generating functions)
is passed as ``hyperparameters'' to \codefun{strucplot}.
This central function 
sets up the graphical layout using grid viewports (see Figure \ref{fig:layout}),
and coordinates the specified core, labeling, shading, and spacing functions to produce 
the plot. On the third level, we provide
several convenience functions such as \codefun{mosaic},
\codefun{sieve}, \codefun{assoc}, and \codefun{doubledecker} which 
interface \codefun{strucplot} through sensible parameter defaults
and support for model formulas. Finally, on the fourth
level, there are `related' \pkg{vcd} functions (such as \codefun{cotabplot}
and the \codefun{pairs} methods for table objects) 
arranging collections of plots of the strucplot 
framework into more complex displays (e.g., by means of panel functions).

>>>>

The strucplot framework is highly modularized: Figure \ref{fig:struc}
shows the hierarchical relationship between the various components. 
On the lowest level, there are several groups of workhorse and 
parameter functions that directly or indirectly influence the final
appearance of the plot. These are examples of `graphical appearance control'
(`grapcon') functions. They are created by generating functions, allowing
flexible parameterization and extensibility. 
The first part of their names (\code{\emph{group\_foo()}}) 
reflects the group they belong to (strucplot core, labeling,
legend, shading, or spacing). The workhorse functions (created by 
\code{\emph{struc\_foo()}}, 
\code{\emph{labeling\_foo()}}, and \code{\emph{legend\_foo()}}) 
directly produce graphical output (``add ink to the canvas''), whereas
the parameter functions
(created by \code{\emph{spacing\_foo()}} and \code{\emph{shading\_foo()}}) compute
graphical parameters used by the others. The grapcon functions returned by 
\code{\emph{struc\_foo()}} implement the core functionality, creating the tiles and their
content.  On the second level of the framework, a suitable combination
of the low-level grapcon functions (or, alternatively, corresponding generating functions)
is passed as ``hyperparameters'' to \codefun{strucplot}.
This central function 
sets up the graphical layout using grid viewports (see Figure \ref{fig:layout}),
and coordinates the specified core, labeling, shading, and spacing functions to produce 
the plot. On the third level, we provide
several convenience functions such as \codefun{mosaic},
\codefun{sieve}, \codefun{assoc}, and \codefun{doubledecker} which 
interface \codefun{strucplot} through sensible parameter defaults
and support for model formulas. Finally, on the fourth
level, there are `related' \pkg{vcd} functions (such as \codefun{cotabplot}
and the \codefun{pairs} methods for table objects) 
arranging collections of plots of the strucplot 
framework into more complex displays (e.g., by means of panel functions).
